Realizability-Preserving Two-Moment DG Method for Fermionic Transport \tnoteref{ldrd-support}

### 2018-01-26 Documents/M1/realizableFermionicM1/sections/realizability.tex
Halfway on Lemma 2.

    To prove $\big(1-\cJ\big)\cJ-|\vect{\cH}|\ge0$ for 3-dimensional $\vect{\cH}$, let's introduce a unit vector $\bn$ with a arbitrary direction.
    Then 
    \begin{equation}
    \vect{\cH} \cdot \bn = |\vect{\cH}| \cos\alpha,
    \end{equation}
    with $\alpha$ be the angle between $\vect{\cH}$ and $\bn$. 
    We know that $-|\vect{\cH}|\leq\vect{\cH} \cdot \bn\leq|\vect{\cH}|$ since $-1\leq\cos\alpha\leq1$.
    While given by definition,
    \begin{align}
    \vect{\cH} \cdot \bn & = \f{1}{4\pi}\int_{\bbS^{2}}f(\omegaNu)\ \bl(\omegaNu) d\omegaNu \cdot \bn \nonumber \\
     &= \f{1}{4\pi}\int_{\bbS^{2}}f(\omegaNu)\ \bl(\omegaNu) \cdot \bn d\omegaNu \nonumber \\
     &= \f{1}{4\pi}\int_{0}^{\pi}\int_{0}^{2\pi}f(\theta,\phi)\ \cos \theta \sin\theta d\theta d\phi
    \end{align}
    while $\theta$ is the angle between $\bl$ and $\bn$, and the coordinate are rotated so that the $\theta$ from $\omegaNu(\theta,\phi)$ is the same with the angle between $\bl$ and $\bn$.
    Note the above expression can also be rewritten by changing variables as
    \begin{align}
    \vect{\cH} \cdot \bn & = \f{1}{2}\int_{0}^{\pi} \f{1}{2\pi}\int_{0}^{2\pi}f(\theta,\phi)\ d\phi \cdot \cos \theta \sin\theta d\theta  \\
    & =  \f{1}{2}\int_{-1}^{1} \f{1}{2\pi}\int_{0}^{2\pi}f(\mu,\phi)\ d\phi \mu d\mu.
    \end{align}
    Besides,
    \begin{align}
    \cJ = \f{1}{4\pi}\int_{\bbS^{2}}f(\omegaNu)\,d\omegaNu = \f{1}{2}\int_{-1}^{1} \f{1}{2\pi}\int_{0}^{2\pi}f(\mu,\phi)\ d\phi d\mu.
    \end{align}
    If we define $f(\mu) =  \f{1}{2\pi}\int_{0}^{2\pi}f(\mu,\phi)\ d\phi$, then
    \begin{align}
    \cJ &= \f{1}{2}\int_{-1}^{1} f(\mu) d\mu, \\
    \vect{\cH} \cdot \bn  &= \f{1}{2}\int_{-1}^{1} f(\mu) \mu d\mu.
    \end{align}
    We know $0\leq\,f(\mu)\leq1$.
    Recall the proof we gave for 1-dimension $\vect{\cH}$, we have
    \begin{equation}
     \big(1-\cJ\big)\cJ-\vect{\cH} \cdot \bn\ge0.
    \end{equation}
    Note this inequality is valid for all the possible $\bn$ which includes the special case with $\cos\alpha=1$ and
    \begin{equation}
    \big(1-\cJ\big)\cJ-|\vect{\cH}|\ge0.
    \end{equation} 
\end{proof}

@@

\begin{lemma}
  The set $\cR$, defined in Eq.~\eqref{eq:realizableSet}, is convex.  
\end{lemma}
\begin{proof}
  Prove this
\end{proof}

>>>>

    To prove $\big(1-\cJ\big)\cJ-|\vect{\cH}|\ge0$ for 3-dimensional $\vect{\cH}$, let's introduce a unit vector $\bn$ with a arbitrary direction.
    Then 
    \begin{equation}
    \vect{\cH} \cdot \bn = |\vect{\cH}| \cos\alpha,
    \end{equation}
    with $\alpha$ be the angle between $\vect{\cH}$ and $\bn$. 
    We know that $-|\vect{\cH}|\leq\vect{\cH} \cdot \bn\leq|\vect{\cH}|$ since $-1\leq\cos\alpha\leq1$.
    While given by definition,
    \begin{align}
    \vect{\cH} \cdot \bn & = \f{1}{4\pi}\int_{\bbS^{2}}f(\omegaNu)\ \bl(\omegaNu) d\omegaNu \cdot \bn \nonumber \\
     &= \f{1}{4\pi}\int_{\bbS^{2}}f(\omegaNu)\ \bl(\omegaNu) \cdot \bn d\omegaNu \nonumber \\
     &= \f{1}{4\pi}\int_{0}^{\pi}\int_{0}^{2\pi}f(\theta,\phi)\ \cos \theta \sin\theta d\theta d\phi
    \end{align}
    while $\theta$ is the angle between $\bl$ and $\bn$, and the coordinate are rotated so that the $\theta$ from $\omegaNu(\theta,\phi)$ is the same with the angle between $\bl$ and $\bn$.
    Note the above expression can also be rewritten by changing variables as
    \begin{align}
    \vect{\cH} \cdot \bn & = \f{1}{2}\int_{0}^{\pi} \f{1}{2\pi}\int_{0}^{2\pi}f(\theta,\phi)\ d\phi \cdot \cos \theta \sin\theta d\theta  \\
    & =  \f{1}{2}\int_{-1}^{1} \f{1}{2\pi}\int_{0}^{2\pi}f(\mu,\phi)\ d\phi \mu d\mu.
    \end{align}
    Besides,
    \begin{align}
    \cJ = \f{1}{4\pi}\int_{\bbS^{2}}f(\omegaNu)\,d\omegaNu = \f{1}{2}\int_{-1}^{1} \f{1}{2\pi}\int_{0}^{2\pi}f(\mu,\phi)\ d\phi d\mu.
    \end{align}
    If we define $f(\mu) =  \f{1}{2\pi}\int_{0}^{2\pi}f(\mu,\phi)\ d\phi$, then
    \begin{align}
    \cJ &= \f{1}{2}\int_{-1}^{1} f(\mu) d\mu, \\
    \vect{\cH} \cdot \bn  &= \f{1}{2}\int_{-1}^{1} f(\mu) \mu d\mu.
    \end{align}
    We know $0\leq\,f(\mu)\leq1$.
    Recall the proof we gave for 1-dimensional $\vect{\cH}$, we have
    \begin{equation}
     \big(1-\cJ\big)\cJ-\vect{\cH} \cdot \bn\ge0.
    \end{equation}
    Note this inequality is valid for all the possible $\bn$ which includes the special case with $\cos\alpha=1$ and
    \begin{equation}
    \big(1-\cJ\big)\cJ-|\vect{\cH}|\ge0.
    \end{equation} 
\end{proof}

@@

\begin{lemma}
  The set $\cR$, defined in Eq.~\eqref{eq:realizableSet}, is convex.  
\end{lemma}
\begin{proof}
  A convex set $Q$ is a set has the following property: 
  \begin{align}
  \theta a + (1-\theta)b \in Q, \text{\space} \text{for } \forall a, b \in Q \text{ and } 0\leq\,\theta\,\leq1. 
  \end{align} 
  Let $\vect{\cM}_{a}=\big(\cJ_{a},\vect{\cH}_{a}\big)^{T}$ and $\vect{\cM}_{b}=\big(\cJ_{b},\vect{\cH}_{b}\big)^{T}$ to be two arbitrary elements of $\cR$.
  If $\cR$ is convex, we will have $\vect{\cM}_{ab} = \big(\cJ_{ab},\vect{\cH}_{ab}\big)^{T} = \theta \vect{\cM}_{a} + (1-\theta)\vect{\cM}_{b}$ in $\cR$ with $0\leq\theta\leq1$.
  Let's prove it is true.

  For the first component
  \begin{align}
  \cJ_{ab} = \theta \cJ_{a} + (1-\theta)\cJ_{b},
  \end{align}
  since $\theta$, $\cJ_{a}$, $1-\theta$ and $\cJ_{b}$ are all nonnegative, $\cJ_{ab}$ is also nonnegative ($\cJ_{ab}\geq0$).
  Meanwhile, since $\cJ_{a}\leq1$ and $\cJ_{b}\leq1$, 
  \begin{align}
  \cJ_{ab} \leq \theta \cdot 1 + (1-\theta) \cdot 1 = 1.
  \end{align}
  Therefore, $\cJ_{ab} \in [0,1]$.

  For the second component,
  \begin{align}
  \vect{\cH}_{ab} 
  & = \theta \vect{\cH}_{a} + (1-\theta)\vect{\cH}_{b} \nonumber\\
  & = \f{1}{4\pi}\int_{\bbS^{2}} \left[ \theta\,f_{a}(\omegaNu) + (1-\theta) f_{b}(\omegaNu) \right] \bl(\omegaNu) d\omegaNu. 
  \end{align}
  Since $\theta\,f_{a}(\omegaNu) + (1-\theta) f_{b}(\omegaNu) \in [0,1]$, we can define $\tilde{f} = \theta\,f_{a}(\omegaNu) + (1-\theta) f_{b}(\omegaNu)$.
  $\tilde{f} \ in [0,1]$ and 
  \begin{align}
  \cJ_{ab} = \f{1}{4\pi}\int_{\bbS^{2}} \tilde{f} d\omegaNu \\
  \vect{\cH}_{ab} = 
  \end{align}
\end{proof}
